Self-tunable approximated explicit MPC: Heat exchanger implementation and analysys

### 2023-02-02 paper/elsarticle/main.tex
schemes
Review: Introduction

	One of the promising control techniques came with the formation of model predictive control~\cite{Morari_MPC}. MPC provides optimal control input based on minimization of a specified cost function, while considering model of the system. Compared to linear quadratic controllers (LQR)~\cite{LQR}, model predictive control also includes constraints on the control input or process variables~\cite{Maciejowski_MPC}, so further saturation is not necessary. Moreover, as the optimization problem is solved in every control step, MPC represents a receding horizon control~\cite{receding_horizon} which has a significant benefit mainly in the terms of disturbance rejection. The model predictive control was intensively investigated in connection with heat exchangers. In~\cite{Vinaya_HE_MPC}, the authors developed a model predictive control for a shell and tube heat exchanger. In~\cite{Oravec_energy}, the robust MPC of a plate heat exchanger with soft constraints was analyzed. Four robust control strategies were presented and compared in~\cite{Oravec_HE_ATE}. A two-level control structure was applied on a heat exchanger network in ~\cite{Gonzales_HE_MPC}. In this study, the low level of control was ensured by MPC and the high level by a supervisory online optimiser.

	The applicability of model predictive control expanded with the parametric solution of MPC optimization problem, known as explicit MPC~\cite{Bemporad_automatica}. As the MPC optimization problem is pre-solved offline, it does not need to be solved in the online (control) phase. Instead, a piece-wise affine (PWA) control law is evaluated to apply the optimal control action. The lower computational complexity makes the explicit MPC more suitable for practical implementation. Nevertheless, the explicit MPC is not tunable in default as the conventional approach from~\cite{Bemporad_automatica} considers the penalty matrices with fixed structure and values.

@@

	The paper~\cite{self_tunable} pushes the idea of tunable explicit MPC further and deals with the issues of practical implementation. The procedure of self-tunable controller technique is presented. The controller's aggressivity is tuned based on the difference between the reference value and the steady state corresponding to the model linearization point. Tuning based on the distance from the steady state represented a way how to compensate the system's nonlinearity.

	This work directly extends our results presented in~\cite{self_tunable}, where the basic principles of the self-tunable approximated explicit MPC were introduced. In this paper, a novel method of tuning parameter setup is introduced. Compared to~\cite{self_tunable}, the tuning method is based on the size of reference step change. Next, the idea of further scaling of the tuning parameter is elaborated. The interval of the tuning parameter is splitted at some value and every part of the interval corresponds to specific operating conditions specified by control engineer. In such way, e.g. the system's asymmetry can be compensated. Last but not least, the proposed tuning technique was implemented in a control of a laboratory plate heat exchanger. This work also provides the control performance evaluation and comparison with using only one non-tunable controller.

	The paper is organized as follows. First, the theoretical backgorunds are presented. The topics of explicit MPC, tunable explicit MPC and an existing self-tunable technique are elaborated. Then, the novel proposed method of self-tunable procedure is explained. Finally, the results of the controller tuning implementation on a heat exchanger are discussed.

>>>>

	One of the promising control strategies addressing all these issues in an optimal way came with the formation and tuning of the model predictive control (MPC), e.g., see~\cite{Morari_MPC}. MPC provides optimal control input based on minimization of a specified cost function, while considering model of the system. Compared to linear quadratic controllers (LQR)~\cite{LQR}, model predictive control also includes constraints on the control input or process variables~\cite{Maciejowski_MPC}, and additional saturation is not necessary. Moreover, as the optimization problem is solved in every control step, MPC represents a receding horizon control policy~\cite{receding_horizon} having a significant benefit mainly in the terms of disturbance rejection. The model predictive control was intensively investigated in connection with heat exchangers. In~\cite{Vinaya_HE_MPC}, the authors developed a model predictive control for a shell and tube heat exchanger. In~\cite{Oravec_energy}, the control performance of the robust MPC implemented for a plate heat exchanger with soft constraints was analyzed. Four robust control strategies were presented and compared in~\cite{Oravec_HE_ATE}. A two-level control structure was applied on a heat exchanger network in~\cite{Gonzales_HE_MPC}. In this study, the low level of control was ensured by MPC and the high level by a supervisory online optimiser.

	The applicability of model predictive control expanded with the parametric solution of MPC optimization problem, known as explicit MPC~\cite{Bemporad_automatica}. As the MPC optimization problem is pre-solved offline, it does not need to be solved in the online phase, i.e., in the real-time control. Instead, a piece-wise affine (PWA) control law is evaluated to apply the optimal control action in each control step. The lower computational complexity makes the explicit MPC more suitable for practical industrial implementation. Nevertheless, the explicit MPC is not tunable in default as the conventional approach in~\cite{Bemporad_automatica} considers the penalty matrices with fixed structure and values.

@@

	The paper~\cite{self_tunable} pushes the idea of tunable explicit MPC further and deals with the issues of practical industrial-oriented implementation. The procedure of self-tunable controller technique is presented. The controller's aggressivity is tuned based on the difference between the reference value and the steady state corresponding to the model linearization point. Tuning based on the distance from the steady-state operating point represented a way on how to compensate the system's nonlinear behaviour.

	This work directly extends our results presented in~\cite{self_tunable}, where the basic principles of the self-tunable approximated explicit MPC were introduced. In this paper, a novel method of a self-tuning parameter setup is introduced. Compared to~\cite{self_tunable}, the self-tuning method is based on the size of reference step change. Moreover, the idea of further scaling of the tuning parameter is elaborated. The interval of the values of the self-tuning parameter is splitted at some certain value, where each part of the interval corresponds to the specific operating conditions defined by control engineer. In such way, e.g., the system's asymmetric behaviour is compensated. Finally, to investigate the benefits of the proposed approach, the proposed self-tuning control policy was implemented to control of a laboratory-scaled counter-current plate heat exchanger. This work provides the control performance evaluation and analysis with using one non-tunable controller.

	The rest of the paper is organized as follows. First, the theoretical backgrounds are presented in Section~\ref{sec:preliminaries}, where the explicit MPC, the approximated tunable explicit MPC, and an existing self-tunable methods are briefly elaborated. Then, the novel proposed method of self-tunable procedure is explained in detail in Section~\ref{sec:methodology}. Finally, the experimental results of the self-tuning controller implementation on a heat exchanger are discussed in Section~\ref{sec:results}, followed by the main conclusions in Section~\ref{sec:conclusion}.
